#Reproducible Research Project 2: NOAA Storm Data Analysis
####kabletown-ux

#Synopsis 

In this project I will analyze NOAA storm data covering meteorological events taking place between 1950 and November 2011. Two questions I will attempt to address are: 1. Across the United States, which types of events are most harmful with respect to population health? 2. Across the United States, which types of events have the greatest economic consequences?

In summary, the answer to first question is that the most harmful event is tornado with respect to human fatalities (5,633 people across the data collection years) and injuries (91,346 people across the data collection years). Additionally, hail has the greatest negative economic impact on crops while tornados have the most significant negative economic effect on property.

#Data Processing
###General Strategy: 
Data for inititial analysis and plotting will be pulled from a 10K record sample (~6MB) of the entire dataset (~536MB).  Once basic analysis, plotting and document layout are completed, the "large" switch will be set to TRUE to enable loading of the entirety of NOAA's storm data file (~536MB).  This is a cacheing strategy borrowed from previous data analysis assignments.  This may prove to be unsuitable for knitr document creation, but, at the very least, it facilitated prototyping all code in a standard R source code module.

###Load Libraries
```{r}
library( ggplot2 )
require( gridExtra )
```

###Set Environment Variables
```{r}
large <- TRUE
download <- FALSE
```
###Get Data
```{r}
## Condidtional download of the data
if ( download ) {
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
    download.file( url, "../data/StormDF.csv.bz2" )
}
```
###Load Data
```{r, cache=TRUE}
# Conditional dataset size: use full, refreshed, data sets...
if ( large ) {
    
    print( "Using FULL version of stormData..." )
    if ( download ) {
        stormData <- read.csv( bzfile( "../data/StormDF.csv.bz2" ) )
    } else {
        stormData <- read.csv( "../data/repdata-data-StormData.csv" )
    }
    
} else { # ...or smaller samples
    
    print( "Using SAMPLED version of stormData..." )
    stormData <- read.csv( "../data/stormDataSample.csv" )
}
```
###Clean Data 
```{r}
## remove superfluous columns (this matters w/ large datasets!)
## deal w/ borked strings
## impute missing values, if necessary
```
###Analyze Data: Human Fatalities & Injuries
```{r}
fatalities <- aggregate( stormData[ "FATALITIES" ], by = stormData[ c( "EVTYPE" ) ], FUN = sum, na.rm = TRUE )
orderedFatalities <- fatalities[ order( -fatalities$FATALITIES ), ]
#substitute "top 10"
orderedFatalities <- head( orderedFatalities, 10 )
#orderedFatalities <- orderedFatalities[ which( orderedFatalities$FATALITIES > 0 ), ]
print( orderedFatalities )

injuries <- aggregate( stormData[ "INJURIES" ], by = stormData[ c( "EVTYPE" ) ], FUN = sum, na.rm = TRUE )
orderedInjuries <- injuries[ order( -injuries$INJURIES ), ]
#substitute "top 10"
orderedInjuries <- head( orderedInjuries, 10 )
#orderedInjuries <- orderedInjuries[ which( orderedInjuries$INJURIES > 0 ), ]
print( orderedInjuries ) 

```
###Analyze Data: Property & Crop Damages
```{r}
cropDamages <- aggregate( stormData[ "CROPDMG" ], by = stormData[ c( "EVTYPE" ) ], FUN = sum, na.rm = TRUE )
orderedCropDamages <- cropDamages[ order( -cropDamages$CROPDMG ), ]
orderedCropDamages <- head( orderedCropDamages, 10 )
print( orderedCropDamages ) 

propertyDamages <- aggregate( stormData[ "PROPDMG" ], by = stormData[ c( "EVTYPE" ) ], FUN = sum, na.rm = TRUE )
orderedPropertyDamages <- propertyDamages[ order( -propertyDamages$PROPDMG ), ]
orderedPropertyDamages <- head( orderedPropertyDamages, 10 )
print( orderedPropertyDamages ) 
```

#Results
###Results: Human Fatalities & Injuries
```{r}
ggplot( data = orderedFatalities, aes( x = orderedFatalities$EVTYPE, y = orderedFatalities$FATALITIES, fill = EVTYPE ) ) + 
    geom_bar( stat = "identity" ) +
    theme( axis.text.x = element_text( angle = 45, hjust = 1 ) ) +
    xlab( "Meteorological Event" ) +
    ylab( "Fatality Count" ) + 
    ggtitle( "Fatality Count, 1950-2011" )

ggplot( orderedInjuries, aes( x = EVTYPE, y = INJURIES, fill = EVTYPE ) ) + 
    geom_bar( stat="identity", position = "dodge" ) + 
    labs( x = "Meteorological Event", y = "Injury Count" ) + 
    labs( title = "Injury Count, 1950-2011" ) +
    theme( axis.text.x = element_text( angle = 45, vjust = 1, hjust=1 ) )
```


###Results: Property & Crop Damages
```{r}

cropsPlot <- ggplot( orderedCropDamages, aes( x = EVTYPE, y = CROPDMG, fill = EVTYPE ) ) + 
    geom_bar( stat="identity", position = "dodge" ) + 
    labs( x = "Type of Event", y = "Crop Damage (Thousands of Dollars?)" ) + 
    labs( title = "Crop Damage, 1950-2011" ) +
    theme( axis.text.x = element_text( angle = 45, vjust = 1, hjust=1 ) )
print( cropsPlot )

propertyPlot <- ggplot( orderedPropertyDamages, aes( x = EVTYPE, y = PROPDMG, fill = EVTYPE ) ) + 
    geom_bar( stat="identity", position = "dodge" ) + 
    labs( x = "Type of Event", y = "Property Damage (Thousands of Dollars?)" ) + 
    labs( title = "Property Damage, 1950-2011" ) +
    theme( axis.text.x = element_text( angle = 45, vjust = 1, hjust=1 ) )
print( propertyPlot )
#can't get good leayout w/o cxrushing...
#grid.arrange( propertyPlot, cropsPlot, heights=3:3, widths=3:3 )

```
