#Reproducible Research Project 2: NOAA Storm Data Analysis
####R. P. Ruiz

#Synopsis: 

In this project I will analyze NOAA storm data covering meteorological events taking place between 1950 and November 2011. Two questions I will attempt to address are: 1. Across the United States, which types of events are most harmful with respect to population health? 2. Across the United States, which types of events have the greatest economic consequences?

In summary, the answer to first question is that the most harmful event is tornado with respect to fatalities (---- thousand people) and flood with respect to injuries (---- thousand people). Additionally, flood has the greatest economic consequences.

#Data Processing (cache = TRUE?)
##General Strategy: Data for inititial analysis and plotting will be pulled from a 10K record sample (~6MB) of the entire dataset (~536MB).  Once basic analysis, plotting and document layout are completed, the "large" switch will be set to TRUE to enable loading of the entirety of NOAA's storm data file (~536MB).  This is a cacheing strategy borrowed from previous data analysis assignments.  This may prove to be unsuitable for knitr document creation, but, at the very least, it facilitated prototyping all code in a standard R source code module.

##Loading Libraries
```{r}
library( ggplot2 )
```

##Setting Environment Variables
```{r}
large <- FALSE
stormData <- NULL
```
##Getting & Loading Data
```{r}
# Conditional data loading: Pull from file...
if ( is.null( stormData ) ) { 
    
    # Conditional dataset size: use full, refreshed, data sets...
    if ( large ) {
        
        print( "Using FULL version of stormData..." )
        stormData <<- read.csv( "../data/repdata-data-StormData.csv" )
        
    } else { # ...or smaller samples
        
        print( "Using TOY version of stormData..." )
        stormData <<- read.csv( "../data/stormDataSample.csv" )
    }
    
} else { # or use cached copy
    
    print( "Using cached versions of stormData..." )
}
```
#Results

The analysis document must have at least one figure containing a plot.

Your analyis must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total.



#PROCESS
Consider writing your report as if it were to be read by a government or municipal manager who might be responsile for preparing for severe weather events and will need to prioritize resources for different types of events. However, there is no need to make any specific recommendations in your report.
